A Decision Procedure for String Constraints with String-Integer Conversion and Flat Regular Constraints

### 2021-05-09 FMCAD2021/4_dec_proc.tex
opt section

%\znjSide{\textbf{Solved}. I think you need to explain the meanings of compatibility.}

Semenov proved that for any function $f$ that is compatible with addition, 
theory $(\mathbb{N},+,f)$ admits QE  and thus the theory is complete and decidable \cite{Semenov84}. 
Exponential functions are compatible with addition functions, and therefore {$\paexp$} is decidable 
by \cite{Semenov84}.
Particularly, in \cite{Point86} Point gave a detailed QE algorithm for $(\mathbb{N},+,2^x)$, where $f(x)=2^x$.
Unfortunately, Point's algorithm is flawed, one problem 
 lies in the discussion to eliminate exponential terms. 
% making the algorithm fail to handle "$a_0<0$" case.
For example, after  eliminating variable $x$ in formula $\exists x. y\le 2^x \wedge x\le 10$,
the expected result should be $y\le 2^{10} =1024$,
but Francoise's algorithm will return a formula equivalent to $y \le 2^6=64$.
% 这里好像还有另一个与Francoise一起的作者

@@

%\znj{\textbf{Solved, missing $\bigvee$ for $\theta_{+\infty}$}.Pls check whether the output is correct or not, my intuition is not correct.} 
% 之前的算法

\subsection{Compared with Francoise's Origin Algorithm}

Francoise's algorithm \cite{Francoise} has four steps,
corresponding to the four steps of our algorithm respectively.
So we will use the names of our steps to refer to them. 
The divisible predicate $n|x$ is not included in his theory,
instead, the division operation $\frac{x}{n}$ where $n\in \mathbb{N}$ is allowed.
The \textbf{Normalization} and \textbf{QE-with-order} steps in two algorithms are similar,
with some minor differences due to the setting of the theory.
The main differences between the two algorithms lies in the following aspects.

In \textbf{QE-exp},
we adopt the same idea to find sufficient conditions
and necessary conditions for the formula to holds.
But our strategy for choosing parameters are easier to understand and use.
A flaw lies in the $a_i<0$ case in Francoise's algorithm 
\footnote{
In the $a_i<0$ case of \textbf{QE-exp-ineq},
Francoise's algorithm assumes falsely $t(y)\ge 0$.
The reason for this mistake might be the confusion of 
two subtraction operators.
The example mentioned in \ref{sec:3.2} shows this case.
},
our \textbf{QE-exp-ineq} corrects this flaw by 
treating $a_i<0$ case similarly to the $a_i>0$ case.

\textbf{QE-linear} in Francoise's algorithm in some sense includes our
\textbf{QE-exp-div}, because divisible predicates is not introduced in his setting.
In this step,
Francoise uses the permutation groups (just like $S_{(n+1)}$ in \textbf{QE-with-order}) and needs recursively renaming of variables $x_i$.
We instead invoke the QE algorithm of PA without referring to permutation groups, 
other improved QE algorithm can also be used to simplify this step.

Francoise's algorithm has more restrictions on the form
of the formula.
Before \textbf{QE-linear} and \textbf{QE-exp} step,
it needs to transform the formula into disjunction normal form (DNF)
and deals with the basic case when input formula $\theta$ is 
a conjunction of atoms. 
It is well known that transforming an arbitrary formula into 
DNF is costly and the length of the formula may increase exponentially.
So our algorithm remove these limitations and assume no special forms of
the formula other than NNF.

@@

\subsection{Complexity Analysis and Case Study}
In this section, we give a complexity analysis of our decision procedure, and provide an example to demonstrate the procedure.
%\subsection{Complexity Analysis}

@@

Other sub-cases are similar. However, when the abstraction parameter becomes larger,
say, $\alpha =\langle 2,2 \rangle$,
the complexity will grow rapidly,
because \textbf{QE-with-order} can not always be omitted and formulae in \textbf{QE-exp} can not be simplified easily.

>>>>

Semenov proved that for any function $f$ that is compatible with addition, 
theory $(\mathbb{N},+,f)$ admits QE  and thus the theory is complete and decidable \cite{Semenov84}. 
Exponential functions are compatible with addition functions, and therefore {$\paexp$} is decidable 
by \cite{Semenov84}.
Particularly, in \cite{Point86} Point gave a detailed QE algorithm for $(\mathbb{N},+,2^x)$, where $f(x)=2^x$.
Unfortunately, Point's algorithm is flawed, one problem 
 lies in the discussion to eliminate exponential terms. 
% making the algorithm fail to handle "$a_0<0$" case.
For example, after  eliminating variable $x$ in formula $\exists x. y\le 2^x \wedge x\le 10$,
the expected result should be $y\le 2^{10} =1024$,
but Point's algorithm will return a formula equivalent to $y \le 2^6=64$.
% 这里好像还有另一个与Francoise一起的作者

@@

\subsection{Compared with Point's Origin Algorithm}

Point's algorithm \cite{Point86} has four steps,
corresponding to the four steps of our algorithm respectively.
So we will use the names of our steps to refer to them. 
The divisible predicate $n|x$ is not included in his theory,
instead, the division operation $\frac{x}{n}$ where $n\in \mathbb{N}$ is allowed.
The \textbf{Normalization} and \textbf{QE-with-order} steps in two algorithms are similar,
with some minor differences due to the setting of the theory.
The main differences between the two algorithms lies in the following aspects.

In \textbf{QE-exp},
we adopt the same idea to find sufficient conditions
and necessary conditions for the formula to holds.
But our strategy for choosing parameters are easier to understand and use.
A flaw lies in the $a_i<0$ case in Point's algorithm 
\footnote{
In the $a_i<0$ case of \textbf{QE-exp-ineq},
Point's algorithm assumes falsely $t(y)\ge 0$.
The reason for this mistake might be the confusion of 
two subtraction operators.
The example mentioned in \ref{sec:3.2} shows this case.
},
our \textbf{QE-exp-ineq} corrects this flaw by 
treating $a_i<0$ case similarly to the $a_i>0$ case.

\textbf{QE-linear} in Point's algorithm in some sense includes our
\textbf{QE-exp-div}, because divisible predicates is not introduced in his setting.
In this step,
Point uses the permutation groups (just like $S_{(n+1)}$ in \textbf{QE-with-order}) and needs recursively renaming of variables $x_i$.
We instead invoke the QE algorithm of PA without referring to permutation groups, 
other improved QE algorithm can also be used to simplify this step.

Point's algorithm has more restrictions on the form
of the formula.
Before \textbf{QE-linear} and \textbf{QE-exp} step,
it needs to transform the formula into disjunction normal form (DNF)
and deals with the basic case when input formula $\theta$ is 
a conjunction of atoms. 
It is well known that transforming an arbitrary formula into 
DNF is costly and the length of the formula may increase exponentially.
So our algorithm remove these limitations and assume no special forms of
the formula other than NNF.

@@

\iffalse
\subsection{Complexity Analysis and Case Study}
In this section, we give a complexity analysis of our decision procedure, and provide an example to demonstrate the procedure.
%\subsection{Complexity Analysis}

@@

Other sub-cases are similar. However, when the abstraction parameter becomes larger,
say, $\alpha =\langle 2,2 \rangle$,
the complexity will grow rapidly,
because \textbf{QE-with-order} can not always be omitted and formulae in \textbf{QE-exp} can not be simplified easily.
\fi

### 2021-05-09 FMCAD2021/5_optimization.tex
pruning opt
update the result of arithmetic
opt section

@@

>>>>

%!TEX root = paper.tex

In the last section, 
we introduced quantifier elimination as a decision procedure for {$\paexp$}. According to theorem \ref{thm:string-parInt}, the satisfiability of {$\strint$} fragment is decidable.
However, the quantifier elimination procedure can not be directly applied on solving string constraints over {$\strint$} due to the following two reasons.
First, quantifier elimination does not return a model that satisfies the constraints. In constrast, it only leaves a formula with constants and free variables. And to obtain satisfiability, all free variables need to be treated as existential quantified variables and be eliminated one by one, 
Second, the quantifier elimination procedure is highly complicated. Note that we alternatively invoke QE-exp and Cooper's algorithm to eliminate exponential terms and linear terms of a quantified variable, the length of the formula grows super-exponentially in the procedure.

In this section, we introduced some optimizations on the QE procedure when focus on the quantifier-free fragment in {$paexp$}. Given a quantifier-free formula, we wish to decide its satisfiability and find an assignment for variables in the formula if it exists. Since the Cooper's algorithm is the main contribution to the complexity, we try to only eliminate exponential terms using an variation of QE-exp and send the obtained linear formula to SMT-solver. The cost is that the algorithm is no longer complete in some situations, in which case some practical restrictions are needed.

A quantifier-free formula is a boolean combination of constraints. Here we allow only equalities and inequalities, divisible predicates are replaced by equalities, for example, $2|x$ is replace to $x=2y$, where $y$ is a fresh variable. Since we only eliminate exponential terms, the newly introduced $y$ will not bother. For convenience, we call variables that occurs exponentially in the formula \emph{exponential variables}, variables that have only linear occurrences \emph{linear variables}. Similarly,
constraints that have exponential terms are called exponential constraints, others are called linear constraints.

\paragraph{QE-exp for all atoms}

\paragraph{over-approximation}

\paragraph{DFS and pruning}

@@

existential PAexp formula, framework, eliminate exponential terms leaving linear terms, send the (equivalent) linear formula to SMT solver

over-approximation for UNSAT and find possible orders

eliminate exponential terms for the formula

dfs search
